### 2024-10-30 Paper_Summary.Rmd
Added Figure 1b. code interpretation from original paper.

---
title: "Paper Summary"
author: "Marqus Odisho & Sargon Yousefian"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggplot2")
#install.packages("lme4")
#install.packages("lmerTest")
#install.packages("MuMIn")
#install.packages("MASS")
#install.packages("dplyr")
#install.packages("tidyr")
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(MASS)
library(dplyr)
library(tidyr)
```

```{r Assay 1: Figure 1a. Simulation & Analysis}
###Trait of Interest: Creating normally distributed female fly populations with average value corresponding to their average number of sexual interactions. Four populations corresponding to the 4 different combinations between female fly size and environment type. (small females in cages, small females in vials, large females in cages, large females in vials)

# Sample Data (replace with your actual data)
data <- data.frame(
  Environment = factor(rep(c("Vial", "Cage"), each = 2), levels = c("Vial", "Cage")),
  Size = factor(rep(c("Small", "Large"), 2), levels = c("Small", "Large")),
  Mean = c(4, 10, 6, 5),
  SD = c(1, 1, 1, 1)  # Standard deviation (replace with your actual values)
)

# Create the bar graph with error bars (Environment on x-axis)
ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), 
                width = 0.2, position = position_dodge(0.9)) +
  labs(
    title = "Random Graph Nonsense",
    x = "",
    y = "Average Number of Sexual Interactions in a 1 Minute Period",
    fill = " " 
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill=NA)
  ) +
  scale_fill_manual(values = c("#B6D0E2", "#F1E3D3"))  # Pastel blue and yellow for the legend
```

```{r Assay 1: Figure 1b. Simulation & Analysis}
MeanFeeding_Small<-c(10,9)
MeanFeeding_Large<-c(6.5,13.5)
EnvironmentType<-c('Vial','Cage')
x<-c(1,2)
SD = c(1, 1, 1, 1)

plot(MeanFeeding_Small, axes=FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim=c(4,16), pch=15, col="#B6D0E2", type='o', lwd=2, cex=1.5)
axis(2)
axis(1, at=seq_along(MeanFeeding_Small), labels = as.character(EnvironmentType), las=0)
lines(c(1,2), MeanFeeding_Large, col="#F1E3D3", pch=17, type="o", lty = 2, lwd=2, cex=1.5)

# Vertical error bars the good ole fashion way
arrows(x0=x, y0=MeanFeeding_Small-SD, x1=x, y1=MeanFeeding_Small+SD, code=3, angle=90, length=0.1)
arrows(x0=x, y0=MeanFeeding_Large-SD, x1=x, y1=MeanFeeding_Large+SD, code=3, angle=90, length=0.1)

# Creating a legend that labels the female size types with different colours - corresponds to the colours in the previous graph
legend('topleft',inset=0.05,c("Small","Large"), lty=c(1,2), col=c("#B6D0E2","#F1E3D3"), lwd=2, title="Female Size")

# Turn plot outline into a box to make it more cutesy
box()
```