### 2024-10-30 Paper_Summary.Rmd

---
title: "Paper Summary"
author: "Marqus Odisho & Sargon Yousefian"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('ggplot2')
#install.packages('lme4')
#install.packages('lmerTest')
#install.packages('MuMIn')
#install.packages('MASS')
#install.packages('dplyr')
#install.packages('tidyr')
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(MASS)
library(dplyr)
library(tidyr)
```

```{r Assay 1: Figure 1a. Simulation & Analysis}
###Trait of Interest: Creating normally distributed female fly populations with average value corresponding to their average number of sexual interactions. Four populations corresponding to the 4 different combinations between female fly size and environment type. (small females in cages, small females in vials, large females in cages, large females in vials)

# Sample Data (replace with your actual data)
data <- data.frame(
  Environment = factor(rep(c('Vial', 'Cage'), each = 2), levels = c('Vial', 'Cage')),
  Size = factor(rep(c('Small', 'Large'), 2), levels = c('Small', 'Large')),
  Mean = c(4, 10, 6, 5),
  SD = c(1, 1, 1, 1)  # Standard deviation (replace with your actual values)
)

# Create the bar graph with error bars (Environment on x-axis)
ggplot(data, aes(x = Environment, y = Mean, fill = Size)) + scale_y_continuous(breaks=c(0, 2, 4, 6, 8, 10)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), 
                width = 0.2, position = position_dodge(0.9)) +
  labs(
    title = "Random Graph Nonsense",
    x = '',
    y = "Average Number of Sexual Interactions in a 1 Minute Period",
    fill = ' ' 
  ) +
  theme_classic() +
  theme(
    legend.position = 'bottom',
    panel.border = element_rect(colour = 'black', fill = NA)
  ) +
  scale_fill_manual(values = c('#B6D0E2', '#F1E3D3'))  # Pastel blue and yellow for the legend
```

```{r Assay 1: Figure 1b. Simulation & Analysis}

#Creating useful variables to create the plot for Figure 1b.
MeanFeeding_Small<-c(10, 9)
MeanFeeding_Large<-c(6.5, 13.5)
EnvironmentType<-c('Vial', 'Cage')
x <- c(1, 2)
SD = c(1, 1, 1, 1)

plot(MeanFeeding_Small, axes=FALSE, xlab= "Mating Arena", ylab="Average Number of Feedings Observed", ylim = c(4, 16), pch = 15, col = '#B6D0E2', type = 'o', lwd = 2, cex = 1.5)
axis(2)
axis(1, at = seq_along(MeanFeeding_Small), labels = as.character(EnvironmentType), las = 0)
lines(c(1, 2), MeanFeeding_Large, col='#F1E3D3', pch = 17, type = 'o', lty = 2, lwd = 2, cex = 1.5)

# Vertical error bars the good ole fashion way
arrows(x0 = x, y0 = MeanFeeding_Small - SD, x1 = x, y1 = MeanFeeding_Small + SD, code = 3, angle = 90, length = 0.1)
arrows(x0 = x, y0 = MeanFeeding_Large - SD, x1 = x, y1 = MeanFeeding_Large + SD, code = 3, angle = 90, length = 0.1)

# Creating a legend that labels the female size types with different colours - corresponds to the colours in the previous graph
legend('topleft', inset = 0.05, c('Small', 'Large'), lty = c(1, 2), col =c ('#B6D0E2','#F1E3D3'), lwd = 2, title = "Female Size")

# Turn plot outline into a box to make it more cutesy
box()
```