### 2024-11-05 Registered Report.Rmd
Starting registered report

---
title: "Registered Report"
author: "Marqus Odisho"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('ggplot2')
#install.packages('lme4')
#install.packages('lmerTest')
#install.packages('MuMIn')
#install.packages('MASS')
#install.packages('dplyr')
#install.packages('tidyr')
#install.packages('car')
#install.packages('pbkrtest')
#install.packages('pscl')
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)
library(MASS)
library(dplyr)
library(tidyr)
library(car)
library(pbkrtest)
library(pscl)
```

```{r Data Simulation}
q_maternal <- rbeta(1000, 0.5, 0.5)
q_paternal <- rbeta(1000, 0.5, 0.5)
plot(density(q_maternal))

# Add correlation test at one point
```